
\zsavepos{header-1}

\zsavepos{header-1}\section{Conclusion and Future work\zsavepos{header-2\zsavepos{header-2}}
\zsavepos{header-3}\zlabel{header}
}\zsavepos{header-3}\zlabel{header}

We propose the \benchmark benchmark, where we systematically study the progress of current procedural graph extraction methods and explore the potential of emerging LLMs on this task. We equip \benchmark with a high-quality dataset, which is about ten times larger than the previous largest ones. Experimental results of baselines in \benchmark reveal that current methods are far from solving this task well. We further involve three advanced LLMs in \benchmark to demonstrate their advantages in extracting textual elements and challenges in organizing local structures. To overcome the main challenge, we design a novel self-refine strategy to empower the LLMs' ability in reasoning gateways. The results show that, with effective strategies, LLMs have the potential of LLMs to comprehend the logical structure among non-sequential actions.

We hope \benchmark can benefit the research on optimal procedural graphs extraction. There are several directions for further work.
First, to improve the performance of LLMs, we suggest introducing procedural knowledge during the pre-training stage of LLMs.
Second, to gain better flow extraction, we will explore more effective methods for handling complex logic structures in the documents.
Lastly, we plan to find a real-world scenario to investigate, besides accuracy, what else is limiting the practical usage of automatic procedure graph extraction.
